---
title: "odd occurence"
author: "Victor Ordu"
date: "1 June 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
_**Caveat: The knowledgable reader will immediately see that the data used for this post is far from ideal (in all respects). The real aim of submitting this post, in spite of the obvious shortcomings, is to help prevent what could be a source of needless frustration to many a data scientist.**_

I was trying to fix up a data entry interface using the `DataEntry` package and in trying to enter the levels of a given factor I came across this happenstance:

```{r}
df <- read.csv("database.csv")
# ... levels of other columns checked
df[[4]]
``` 

When I saw this output, I figured that in order for me to see the levels accurately I would need to call `levels()`. However, since I was trying to do a quick job, I wanted to copy the levels to the clipboard and paste into the appropriate field in `DataEntry`. So I didn't need all the other stuff that would come with regular printing to the console and I chose to use `cat`
```{r}
cat(levels(df[[4]]))
```

Which looks pretty good and worked just fine in terms of convenience. But then I decided I needed to also view a regular printout of the levels since they were so jumbled up.

```{r}
levels(df[[4]])
```

I was shocked! I had almost totally missed the first element among the given categories because it contained NOTHING!

The horror this would have unleashed on my further work with the data is best imagined (after all, I was just trying to draft a data entry format and fix the factors later).

There are so many lessons to be learnt from this near-fiasco.

Tread carefully.